---
title: 'Learn R with Addiction Science: Group Practice Questions - Set 2'
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

First, let's load our data, and the TidyVerse.

```{r}
# Only use these if you do not have them installed:
# install.packages("tidyverse")
# intstall.packages("palmerpenguins")
library(tidyverse)
library(nycflights13)

# Now let's save our data as an object
df <- flights
View(flights)

# And let's inspect what else is in the package
?airlines
?airports
?planes
?weather
```

For these practice questions, we'll be using all of the functions we've practiced so far:

-   ggplot() + all of the different types of graphs (or "geoms")

    -   scatter plot: geom_point()

    -   boxplot: geom_box()

    -   bar chart: geom_bar()

-   filter() - Filter in/out data that you are not interested in. Can be used to "subset", or split a larger data set into smaller data sets based on criteria.

-   summarize() - Can be used to create smaller summaries of a data set.

    -   mean() - Can be used to calculate the mean of a column, list, or other collection of numbers

    -   sd() - Can be used to calculate the standard deviation of a column, list, or other collection of numbers

    -   All common central tendency statistics (mean, median, mode) and measures of spread (standard deviation, variance, etc) have similar functions.

-   group_by() - Can use one column to group the data by. Then, can be used to feed into summaries to group our summarized data

-   select() - With this function, we can make a smaller dataset by selecting only the columns we want.

-   mutate() - This function will help you create new variables or modify existing ones.

-   as_factor() - Converts data from a different type to a factor (categorical) type.

Now that we've reviewed these functions, use them to answer the questions below:

1.  What is the average speed of every flight by month?

    ```{r}
df <- df %>% mutate(flight_speed_mph = distance/air_time*60)
df %>% group_by(month) %>% summarise(mean(flight_speed_mph, na.rm = TRUE))
    ```

2.  Create a bar graph of the number of flights per carrier. Which carrier (name, not abbreviation) has the most flights?

    ```{r}

ggplot(data = flights, mapping = aes(y = carrier))+geom_bar()
    ```

3.  What is the average delay in arrival for planes originating from JFK in March?

    ```{r}
df %>% filter(month == 3, origin == "JFK") 
    
  df %>% summarise(mean(arr_delay, na.rm = TRUE))
    ```

4.  Create a scatterplot which shows the relationship between the delay in departure and the delay in arrival. What kind of relationship do these numbers have?

    ```{r}
ggplot(data = flights, mapping = aes(x = dep_delay, y = arr_delay))+
      geom_point()
    
#Positive correlation ~1. As departure delay increases, so does arrival delay.
    ```

5.  What is the average distance traveled by originating airport in kilometers?

    ```{r}

    ```
